### 2023-09-01 analysis/2_registered.Rmd
Summary stats for key measures

---
title: "Visual Search and Pre-crastination"
author: "Alasdair D.F. Clarke"
date: '2022-05-30'
output:
  tufte::tufte_html:
    number_sections: true
    fig_height: 3
    link-citations: yes  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits = 3)
set.seed(2022)
```

```{r, message=FALSE}
# attach packages
#library(rethinking)
library(tidyverse)
library(patchwork)
library(binom)
```

```{r}
# set ggplot2 options
theme_set(ggthemes::theme_tufte() + 
            theme(plot.background = element_rect(fill = "#fffff8", 
                                                 colour = "#fffff8")))

our_cols <- c("#1b9e77", "#7570b3")
options(ggplot2.discrete.fill = our_cols ,
        ggplot2.discrete.colour = our_cols)
```

# Import data

```{r}
dat_s <- read_csv("scratch/dat_rt_acc_strat.csv")

dat_p <- read_csv("scratch/prcrastination_and_personality.csv") %>%
  filter(!is.na(pre_crastinate)) %>%
  mutate(pre_crastinate = as_factor(pre_crastinate),
         pre_crastinate = fct_relevel(pre_crastinate, "no", "intermediate", "yes"))

```

# Overview of measures

## Visual Search Accuracy & Response Time

```{r}
# merge search with personality and precrastination data
dat <- left_join(dat_s, dat_p) %>%
  filter(!is.na(pre_crastinate))

dat %>% ggplot(aes(targ_side_rel, accuracy, fill = pre_crastinate)) +
  geom_boxplot(alpha = 0.5) +
    scale_x_discrete("trial type")  +
  theme(legend.justification=c(0,0), legend.position=c(0,0))  -> plt_acc

dat %>% ggplot(aes(targ_side_rel, median_rt, fill = pre_crastinate)) +
  geom_boxplot(alpha = 0.5) +
  theme(legend.position = "none") + 
  scale_x_discrete("trial type")  +
  scale_y_log10("reaction time (seconds)")  -> plt_rt

plt_acc + plt_rt

ggsave("scratch/acc_main.pdf", plt_acc + theme_bw() + theme(legend.position = "none"), width = 4, height = 3)
ggsave("scratch/rt_main.pdf", plt_rt  + theme_bw() + theme(legend.justification=c(1,0), legend.position=c(1,0)), width = 4, height = 3)

```

```{r}
rm(plt_acc, plt_rt)
```

## Search Strategy & Pre-crastination

```{r}
dat_s %>% 
  group_by(person) %>%
  summarise(search_strat = unique(prop_hard)) -> dat_strat

dat_strat %>%
  ggplot(aes(search_strat)) + 
  geom_histogram(breaks = seq(0.0, 1, 0.1), colour = "black") + 
  scale_x_continuous("search strategy score") -> plt

ggsave("scratch/strat_hist.pdf", plt + theme_bw() + theme(legend.position = "none"), width = 4, height = 3)

```

And now look at pre-crastination measure.

```{r}

dat_p %>% ggplot(aes(x = n_pc)) +
  geom_histogram(bins = 7, alpha = 0.67, colour = "black") +
  scale_x_continuous("number of trials") + 
  scale_fill_viridis_d() +
  theme(legend.position = "none") -> plt1

dat_p %>% ggplot(aes(x = pre_crastinate)) + 
  geom_bar() +
  scale_x_discrete("pre-crastination classification") + 
  scale_fill_viridis_d() +
  theme(legend.position = "none") -> plt2

plt1 + plt2

ggsave("scratch/pre_crastinate.pdf", plt2 + theme_bw(),  width = 4, height = 3)

```

We have replicated the original pre-crastination finding....

```{r}
n_row <- nrow(dat_p)

dat_p %>% 
  group_by(pre_crastinate) %>%
  summarise(n = n(),
            percentage = 100*n/n_row) %>%
  knitr::kable()

rm(n_row)
```