### 2023-09-01 analysis/3_exploratory.Rmd
Update 3_exploratory.Rmd

---
title: "Visual Search and Pre-crastination"
author: "Alasdair D.F. Clarke"
date: '2022-05-30'
output:
  tufte::tufte_html:
    number_sections: true
    fig_height: 3
    link-citations: yes  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits = 3)
set.seed(2022)
```

```{r, message=FALSE}
# attach packages
library(rethinking)
library(tidyverse)
library(patchwork)

```

```{r}
# set ggplot2 options
theme_set(ggthemes::theme_tufte() + 
            theme(plot.background = element_rect(fill = "#fffff8", 
                                                 colour = "#fffff8")))

```

# Import data

```{r}
dat_s <- read_csv("scratch/dat_rt_acc_strat.csv")

dat_p <- read_csv("scratch/prcrastination_and_personality.csv") %>%
  filter(!is.na(pre_crastinate)) %>%
  mutate(pre_crastinate = as_factor(pre_crastinate),
         pre_crastinate = fct_relevel(pre_crastinate, "no", "intermediate", "yes"))

dat <- full_join(dat_s %>% filter(targ_side_rel == "absent"), 
          dat_p, 
          by = c("person", "site"))
```

# Calculate Correlations

```{r}
measures <- factor(1:10, labels = c("BIStotal", "attentional_imp", "motor_imp", "nonplan_imp",
              "open_mindedness", "conscientiousness", "extraversion", "agreeableness", "negative_emo",
              "procrastination"))

measures <- fct_rev(measures)

dps <- dat %>% filter(pre_crastinate != "intermediate")

get_rd_int <- function(measure) {

  r <- cor.test(dat$prop_hard, dat[[measure]])

  d <- t.test(dps[[measure]] ~ dps$pre_crastinate)

  d_out <- tibble(measure = measure,
                  lower = c(r$conf.int[[1]], d$conf.int[[1]]),
                  upper = c(r$conf.int[[2]], d$conf.int[[2]]),
                  feature = c("search", "pre-crastination"))
  return(d_out)
}

dr <- map_df(measures, get_rd_int)

dr$col <- as_factor(rep(c(1,1,1,1,2,2,2,2,2,3), each = 2))

ggplot(dr, aes(xmin = lower, xmax = upper, y = measure, colour = col)) +
  geom_linerange(linewidth = 2, alpha = 0.7) +
  geom_vline(xintercept = 0, linetype = 2) + 
  scale_x_continuous("effect size") +
  ggthemes::scale_color_fivethirtyeight() +
  theme(legend.position = "none") +
  facet_wrap(~feature)  -> plt

plt

ggsave("scratch/exp_strat.pdf", plt + theme_bw()+
  theme(legend.position = "none"), width = 6, height = 3)
ggsave("scratch/exp_strat.png", plt + theme_bw()+
  theme(legend.position = "none"), width = 6, height = 3)
```